Demystifying the Computer Science PhD Admission in the US\\{\large A Guide for International Students

### 2023-11-24 demystify.tex
updates

\begin{commentbox}[Vu:]
  Sometime PhD admission committee in the US will share a document such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Foreign-Top-Schools}{this one}, which lists the top schools in several countries. In some cases we ask other faculty and students if we think they know about the place.  For example, when I was a postdoc at UMD, people in admission committee ask me to evaluate applicants from Vietnam.  During my time at UNL and now here at GMU, I have looked at Vietnamese applications (whether they are assigned to me or not) and provide input to their reviewers, e.g., X is the top tech school in Vietnam and so it should be \emph{top} instead of \emph{unknown} foreign, which makes a huge difference.
\end{commentbox}
\begin{commentbox}[Deepak:] 
If an applicant is anxious about their school not being known outside their country, they can provide information about their school and department, with independent sources where such informationa can be verified.
\end{commentbox}

@@

This can be an issue for students in many top international universities where the competition is so high that very good students students can still have low rankings from these schools (and be overlooked by the admission committee).
So what to do with this? well, same as \hyperref[sec:your-school]{before}, e.g., put a note about this in your statement and ask your LoR writers to talk about it.

@@

Another effective way is taking graduate level courses in the topics you are interested in (remember: you have those 2 years to explore).  Professors teaching these \emph{special topics courses} and \emph{research seminars} typically are recruiting students---giving you even a higher chance. Do well in the class, answer questions, talk to the profs after classes, etc---being stand out.  Many professors, including myself, prefer taking in new students this way.  It gives both the professor and student more time, e.g., a whole semester, to work and evaluate the relationship before making any commitment (sounds a bit like a marriage!).
You can also ask if you can do an independent study or research with a prof. This can be informal (no credit) and takes place during the summer or winter break.  For example, I work with multiple students, many of whom are undergrads, this way. Many will drop because they find they don't like my research, but some will stay and become my PhD students.

@@

  In other fields, e.g. Software Engineering, Vietnamese students face challenges in reaching US professors. Do you have any tips for Vietnamese students who want to connect with US professors and work as research assistants?
  \tcblower
  \textbf{Vu}: \S\ref{sec:contact} shows how to contact a professor for research opportunities. Many probably will say no as they do not have the bandwidth to take on random students, but some may say yes if they see potential fit.
\end{commentbox}

@@

\begin{commentbox}[DK:]
I was amused to read this as if I recall correctly, you never called me by my first name when you were at UNM. You always called me Prof. And, many times, I would jokingly call you back as Prof. Vu.
\end{commentbox}

@@

There's a misconception that in the US it's all business, with professors as bosses who pay students for their work, and you doing something nice implying you expect something in return. However, the reality is quite the opposite. While people can be straightforward and direct, they are also friendly and informal. You can call your professor by their first name (\S\ref{sec:address}), disagree with them and argue (and can actually gain respect doing so), seek their help (even on personal matters), come to their houses for parties or gathering, and give them small thoughtful gifts that they put on their desks.  Many students and professors maintain lifelong relationships, staying in touch through cards, emails, and calls, even after their academic journey ends.% These connections are genuine and meaningful.

>>>>

\begin{commentbox}[Vu:]
  Sometime PhD admission committee in the US will share a document such as \href{https://github.com/dynaroars/dynaroars.github.io/wiki/Foreign-Top-Schools}{this one}, which lists the top schools in several countries. In some cases we ask other faculty and students if we think they know about the place.  For example, when I was a postdoc at UMD, people in admission committee ask me to evaluate applicants from Vietnam.  During my time at UNL and now here at GMU, I have looked at Vietnamese applications (whether they are assigned to me or not) and provide input to their reviewers, e.g., X is the top tech school in Vietnam and so it should be \emph{top} instead of \emph{unknown} foreign, which makes a huge difference.
\end{commentbox}
\begin{commentbox}[Deepak:] 
If an applicant is anxious about their school not being known outside their country, they can provide information about their school and department, with independent sources where such information can be verified.
\end{commentbox}

@@

This can be an issue for students in many top international universities where the competition is so high that very good students students can still have low rankings from these schools (and be overlooked by the admission committee).
So what to do with this? as mentioned in \S\ref{sec:your-school} you can put a note about this in your statement and ask your LoR writers to talk about it.

@@

Another effective way is taking graduate level courses in the topics you are interested in (remember: you have those 2 years to explore).  Professors teaching these \emph{special topics courses} and \emph{research seminars} might be recruiting students---giving you even a higher chance. Do well in the class, answer questions, talk to the prof after classes, etc---being stand out.  Many professors, including myself, prefer taking in new students this way.  It gives both the professor and student more time, e.g., a whole semester, to work and evaluate the relationship before making any commitment (sounds a bit like a marriage!).
You can also ask if you can do an independent study or research with a prof. This can be informal (no credit) and takes place during the summer or winter break.  For example, I do this with multiple students, many of whom are undergrads. Many will drop because they find they don't like my research, but some will stay and become my PhD students.

@@

  In other fields, e.g. Software Engineering, Vietnamese students face challenges in reaching US professors. Do you have any tips for Vietnamese students who want to connect with US professors and work as research assistants?
  \tcblower
  \textbf{Vu}: \S\ref{sec:contact} shows how to contact a professor for research opportunities. Many probably will say no (or do not reply) as they do not have the bandwidth to take on random students, but some may say yes if they see potential fit.
\end{commentbox}

@@

\begin{commentbox}[DK:]
  I was amused to read this as if I recall correctly, you never called me by my first name when you were at UNM. You always called me Prof. And, many times, I would jokingly call you back as Prof. Vu.
  \tcblower
  \textbf{Vu}: Yes, for some reason I enjoy addressing you as ``Prof.'' (without appending a last or first name).  I call Steph and everyone else by their their first names.  The use of Prof. Vu may have foreshadowed my future in academia.
\end{commentbox}

Certainly, there's a peculiar satisfaction in addressing you as "Prof." without appending a last or first name. While I refer to Steph and everyone else by their first names, perhaps addressing Prof. Vu in this manner hinted at my future in academia.

@@

There's a misconception that in the US it's all business, with professors as bosses who pay students for their work, and you doing something nice implying you expect something in return. However, the reality is quite the opposite. While people can be straightforward and direct, they are also friendly and informal. You can call your professor by their first name (\S\ref{sec:address}), disagree with them and argue (and gain respect doing so), seek their help (even on personal matters), come to their houses for parties or gathering, and give them small thoughtful gifts that they put on their desks.  Many students and professors maintain lifelong relationships, staying in touch through cards, emails, and calls, even after their academic journey ends.% These connections are genuine and meaningful.
